Codebook

### 2019-05-22 reproducing_registered_reports.Rmd
add codebook

abstract: |
  Ongoing technological developments have made it easier than ever before for scientists to share their data, materials, and analysis code. Sharing data and analysis code makes it easier for other researchers to re-use or check published research. These benefits will only emerge if researchers can reproduce the analysis reported in published articles, and if data is annotated well enough so that it is clear what all variables mean. Because most researchers have not been trained in computational reproducibility, it is important to evaluate current practices to identify practices that can be improved. We examined data and code sharing, as well as computational reproducibility of the main results without contacting the original authors, for Registered Reports published in the in psychological literature between 2014 and 2018. Of the 62 articles that met our inclusion criteria data was available for 40 articles, and analysis scripts for 43 articles. For the 35 articles that shared both data and code and performed analyses in SPSS, R, or JASP, we could run the scripts for 30 articles, and reproduce the main results for 19 articles. Although the percentage of articles that shared both data and code (61%) and articles that could be computationally reproduced (54%) was relatively high compared to other studies, there is clear room for improvement. We provide practices recommendations based on our observations, and link to examples of good research practices in the papers we reproduced.

@@

# > sessionInfo()
# R version 3.5.3 (2019-03-11)
# Platform: x86_64-w64-mingw32/x64 (64-bit)
# Running under: Windows 10 x64 (build 17134)
# 
# Matrix products: default
# 
# locale:
# [1] LC_COLLATE=Dutch_Netherlands.1252  LC_CTYPE=Dutch_Netherlands.1252   
# [3] LC_MONETARY=Dutch_Netherlands.1252 LC_NUMERIC=C                      
# [5] LC_TIME=Dutch_Netherlands.1252    
# 
# attached base packages:
# [1] stats     graphics  grDevices utils     datasets  methods   base     
# 
# other attached packages:
# [1] here_0.1       irr_0.84.1     lpSolve_5.6.13 readxl_1.3.1  
# 
# loaded via a namespace (and not attached):
#  [1] Rcpp_1.0.1       crayon_1.3.4     rprojroot_1.3-2  cellranger_1.1.0 backports_1.1.4 
#  [6] pillar_1.4.0     rlang_0.3.4      rstudioapi_0.10  tools_3.5.3      xfun_0.6        
#  [11] compiler_3.5.3   pkgconfig_2.0.2  knitr_1.22       tibble_2.1.1

library(readxl)
library(here)
library(irr)

all_data <- read_excel("Data_for_Analysis_of_Open_Data_and_Computational_Reproducibility_in_Registered_Reports_in_Psychology.xlsx", n_max = 62)

@@

Researchers are currently exploring ways to make science more open and transparent. Among novel developments such as pre-registration, pre-prints, and open peer review, an increasing number of journals, funders, and researchers are beginning to expect that data, materials, and analysis code will be shared by default with scientific publications (e.g., [@morey_peer_2016]. Sharing data and analysis code with scientific publications allows others to more easily reproduce, check, and build on existing work. However, since most researchers have received relatively little training in how to share data and code to facilitate reproducible science, it is important to evaluate how data and code is currently being shared to learn what needs to be improved. With this goal in mind we aimed to computationally reproduce the results of Registered Reports published in the psychological literature.

It is desirable that research is reproducible. Data availability has the potential to save time in the long run by facilitating other researchers to re-use data, allows findings to be more easily included in meta-analyses, and makes it possible for peers to check and correct published findings. According to @kitzes_practice_2017 computational reproducibility means that ”a second investigator (including the original researcher in the future) can recreate the final reported results of the project, including key quantitative findings, tables, and figures, given only a set of files and written instructions.” For scientific research to be computationally reproducible, the data and code needs to be shared. However, the availability of data and code in itself is not enough. Data and code also need to be linked in the article so that readers know where to find it. Preferably, the data should be available in a format that can be read by open source software. In addition to being able to access the data and code underlying the results in an article, researchers should be able to understand the data and code (which means variables are described, and code is annotated). Finally, the results reported in the scientific manuscript should be reproducible, which means the data and code can be used to compute the results that are reported in the article.

Recently, scholars have started to empirically examine the extent to which data is shared with published articles, and if so, whether it was possible to reproduce the data analysis reported in the published article. @hardwicke_data_2018 examined the reproducibility of 35 articles published in the journal Cognition. They could reproduce 11 articles without assistance, 11 with assistance from the original authors, and 13 articles contained at least one outcome that could not be reproduced even with author assistance. Hardwicke et al. (2018) estimated that it took between 2-4 hours to reproduce a paper, but they did not record the exact time. @stockemer_data_2018 analyzed reproducibility in all articles published in 2015 in three political science journals. They emailed authors for the code and data, and found that for 71 articles for which they received code and data, one could not be reproduced due to a lack of a software license, and 16 articles could not be reproduced with access to the required software. For the remaining articles, 32 could be exactly reproduced, 19 could be reproduced with slight differences, and 3 articles yielded significantly different results. @stodden_empirical_2018 analyzed data availability in the journal science in 2011-2012 and found that 26 of 204 (or 13%) of articles provided information to retrieve data and/or code without contacting the authors. For all datasets they acquired after e-mailing authors for data and code, 26% was estimated to be computationally reproducible.

Our main interest is to examine whether data and code authors shared with their paper were reusable without contacting the original authors. The benefit of sharing data and code over making them available upon request is that even if the original author can no longer be reached, the data is still reusable. We set out to examine the data availability and reproducibility in Registered Reports published in psychological science. Registered Reports are a relatively novel development in psychology. The theoretical introduction and a detailed description of the method is submitted for publication, peer reviewed, and revised, after which it can receive an 'in principle acceptance'. This means the article will be published as long as the authors follow their preregistered data collection and analysis plan [@chambers_instead_2014; @nosek_registered_2014]. The population of Registered Reports in psychology is still relatively small (91 Registered Reports had been published as of February 22, 2018), which makes it possible to examine it in full. The novelty of Registered Reports may attract early adopters who are also exploring other novel developments in psychology, such as sharing data and code. We assumed that researchers who publish Registered Reports would also be likely to share data and code in public repositories. In line with our assumptions, `r sum(all_data$availability, na.rm = TRUE)` out of `r length(all_data$doi)` articles in our final dataset shared data and code (`r round(100*sum(all_data$availability, na.rm = TRUE)/length(all_data$doi),1)`%). In comparison, after Cognition introduced a mandatory data sharing policy, 136 out of 174 articles (78.2%) had a data availability statement, and 49% of 174 articles had reusable data (Hardwicke et al., 2018). Sharing analysis code was not required, and only 18 out of 174 articles provided the analysis code (10.3%). To evaluate the reproducibility of findings published in Registered Reports, we examine if data can be located, is available at the indicated location, can be opened in open-source software, is understandable, and can be used to reproduce the reported results.

@@

Our main objective is to examine how reusable data and code underlying Registered Reports in Psychology are, based purely on the information provided in the published literature. We examine the extent to which we can reproduce reported analyses without contacting the original authors. While reproducing the results reported in Registered Reports, we kept track of factors that facilitated reproducibility or that made reproducing the result more difficult. We report these qualitative findings with an aim to highlight how current practices can be improved.

@@

To find Registered Reports published in psychology, we drew from a database of registered reports maintained by by Center for Open Science. At the start of this project (06-05-2018) this database consisted of 118 published Registered Reports. Eighty-one articles in this database were published in the psychological literature. We limited our analysis to studies performed by single groups, since such papers are most representative of the psychological literature, and excluded 7 large scale collaborations where dedicated team members worked on making the analysis reproducible, such as Registered Replication Reports [@hagger_multilab_2016] and Many Labs projects [@klein_investigating_2014; @klein_many_2018]. Upon further inspection of the 72 papers left in the dataset, it turned out that 10 were not formally registered reports. This left `r length(all_data$doi)` studies in our sample. When evaluating whether we could reproduce the original results, we limited ourselves to statistical software packages that we had experience with (R, SPSS, and JASP) and excluded studies that required expertise in software packages we are not trained in (e.g., dedicated EEG software, Python, Matlab), which led to the exclusion of `r sum(all_data$included_language == FALSE, na.rm = TRUE)` additional studies.

@@

We used a dichotomous coding scheme to when coding whether the script could be executed whether the results could be reproduced ('yes' or 'no'). Code often needed minor adjustments to run on the data, such as changing file locations, or loading packages in R, and coders sometimes took different approaches to how much they would adjust the code to make it run on the data. Furthermore, coders initially used different thresholds of reproducibility, based on whether every single result reported in a paper could be reproduced by the code and data, or only the main results reported in the article. After evaluating our initial coding round, we considered an article reproducible if we could get the same main results as represented in the paper with at best minor changes to make the scripts run on the data. This means we considered the analysis reproducible, even if variables could only be found with reasonable effort or by making assumptions about how variable names mapped onto variables reported in the data analysis. Furthermore, we changed folder locations when needed and loaded libraries that were required.

>>>>

abstract: |
  Ongoing technological developments have made it easier than ever before for scientists to share their data, materials, and analysis code. Sharing data and analysis code makes it easier for other researchers to re-use or check published research. These benefits will only emerge if researchers can reproduce the analysis reported in published articles, and if data is annotated well enough so that it is clear what all variables mean. Because most researchers have not been trained in computational reproducibility, it is important to evaluate current practices to identify practices that can be improved. We examined data and code sharing, as well as computational reproducibility of the main results without contacting the original authors, for Registered Reports published in the in psychological literature between 2014 and 2018. Of the 62 articles that met our inclusion criteria data was available for 40 articles, and analysis scripts for 43 articles. For the 35 articles that shared both data and code and performed analyses in SPSS, R, or JASP, we could run the scripts for 30 articles, and reproduce the main results for 19 articles. Although the percentage of articles that shared both data and code (61%) and articles that could be computationally reproduced (54%) was relatively high compared to other studies, there is clear room for improvement. We provide practical recommendations based on our observations, and link to examples of good research practices in the papers we reproduced.

@@

library(readxl)
library(here)
library(irr)
library(codebook)

all_data <- read_excel("Data_for_Analysis_of_Open_Data_and_Computational_Reproducibility_in_Registered_Reports_in_Psychology.xlsx", n_max = 62)
codebook_data <- read_excel("Data_for_Analysis_of_Open_Data_and_Computational_Reproducibility_in_Registered_Reports_in_Psychology.xlsx", sheet = "Codebook")

#Use the codebook package to add labels to variables
var_label(all_data) <- dict_to_list(codebook_data)

@@

Researchers are currently exploring ways to make science more open and transparent. Among novel developments such as pre-registration, pre-prints, and open peer review, an increasing number of journals, funders, and researchers are beginning to expect that data, materials, and analysis code will be shared by default with scientific publications (e.g., [@morey_peer_2016]. Sharing data and analysis code with scientific publications allows others to more easily reproduce, check, and build on existing work. However, since most researchers have received limited training in how to share data and code to facilitate reproducible science, it is important to evaluate how data and code are currently being shared to learn what needs to be improved. With this goal in mind, we aimed to computationally reproduce the results of Registered Reports published in the psychological literature.

It is desirable that research is reproducible. Data availability has the potential to save time in the long run by facilitating the re-use of data, allows findings to be more easily included in meta-analyses, and makes it possible for peers to check and correct published findings. According to @kitzes_practice_2017, computational reproducibility means that ”a second investigator (including the original researcher in the future) can recreate the final reported results of the project, including key quantitative findings, tables, and figures, given only a set of files and written instructions.” For scientific research to be computationally reproducible, the data and code needs to be shared. However, the availability of data and code in itself is not enough. These materials also need to be linked in the article so that readers know where to find them. Preferably, the data should be available in a format that can be read by open source software. Having access to rdata and analysis files is not worth much of these files can not be understood. Variables must be described and labeled (e.g., in a codebook), and code should be annotated. Finally, the results reported in the scientific manuscript should be reproducible, which means the data and code can be used to compute the results that are reported in the published article.

Recently, scholars have started to empirically examine the extent to which data is shared with published articles, and, if so, whether it was possible to reproduce the data analysis reported in the published article. @hardwicke_data_2018 examined the reproducibility of 35 articles published in the journal Cognition. Eleven  articles could be reproduced without assistance from the original authors, and 13 articles contained at least one outcome that could not be reproduced even with author assistance. Hardwicke et al. (2018) estimated that it took between 2-4 hours to reproduce a paper, but they did not record the exact time. @stockemer_data_2018 analyzed reproducibility in all articles published in 2015 in three political science journals. They emailed authors for the code and data, and found that for 71 articles for which they received code and data, one could not be reproduced due to a lack of a software license, and 16 articles could not be reproduced with access to the required software. For the remaining articles, 32 could be exactly reproduced, 19 could be reproduced with slight differences, and 3 articles yielded significantly different results. @stodden_empirical_2018 analyzed data availability in the journal science in 2011-2012 and found that 26 of 204 (or 13%) of articles provided information to retrieve data and/or code without contacting the authors. For all datasets they acquired after e-mailing authors for data and code, 26% was estimated to be computationally reproducible. It seems there is clear room for improvement in how reproducible published articles are.

We set out to examine the data availability and reproducibility in Registered Reports published in psychological science. Our main interest is to examine whether data and code authors shared with their paper were reusable without contacting the original authors. The benefit of sharing data and code over making them available upon request is that even if the original author can no longer be reached, the data is still reusable. Registered Reports are a relatively novel development in psychology. Before data collection commences, the introduction and methods are peer-reviewed, after whichauthors can receive an 'in principle acceptance'. This means the article will be published as long as the authors follow their preregistered data collection and analysis plan [@chambers_instead_2014; @nosek_registered_2014]. The population of Registered Reports in psychology is still relatively small (91 Registered Reports had been published as of February 22, 2018), which makes it possible to examine it in full. 
The novelty of Registered Reports may attract early adopters who are also exploring other novel developments in psychology, such as sharing data and code. We assumed that researchers who publish Registered Reports would also be likely to share data and code in public repositories. In line with our assumptions, `r sum(all_data$availability, na.rm = TRUE)` out of `r length(all_data$doi)` articles in our final dataset shared data and code (`r round(100*sum(all_data$availability, na.rm = TRUE)/length(all_data$doi),1)`%). In comparison, after *Cognition* introduced a mandatory data sharing policy, 136 out of 174 articles (78.2%) had a data availability statement, and 49% of 174 articles had reusable data (Hardwicke et al., 2018). Sharing analysis code was not required, and only 18 out of 174 articles provided the analysis code (10.3%). To evaluate the reproducibility of findings published in Registered Reports, we examine if data can be located, are available at the indicated location, can be opened in open-source software, are understandable, and can be used to reproduce the reported results.

@@

Our main objective was to examine how reusable data and code underlying Registered Reports are, based purely on the information provided in the published literature. We examined the extent to which we could reproduce reported analyses without contacting the original authors. While reproducing the results reported in Registered Reports, we kept track of factors that facilitated reproducibility or that made reproducing the result more difficult. We report these qualitative findings with an aim to highlight how current practices can be improved.

@@

To find Registered Reports published in psychology, we drew from a database of registered reports maintained by by Center for Open Science. At the start of this project (06-05-2018) this database consisted of 118 published Registered Reports. Eighty-one articles in this database were published in the psychological literature. We limited our analysis to studies performed by single groups, since such papers are most representative of the psychological literature, and excluded 7 large scale collaborations where dedicated team members worked on making the analysis reproducible, such as Registered Replication Reports [@hagger_multilab_2016] and Many Labs projects [@klein_investigating_2014; @klein_many_2018]. Upon further inspection of the 72 papers left in the dataset, it turned out that 10 were not formally Registered Reports. This left `r length(all_data$doi)` studies in our sample. When evaluating whether we could reproduce the original results, we limited ourselves to statistical software packages that we had experience with (R, SPSS, and JASP) and excluded studies that required expertise in software packages we are not trained in (e.g., dedicated EEG software, Python, Matlab), which led to the exclusion of `r sum(all_data$included_language == FALSE, na.rm = TRUE)` additional studies.

@@

We used a dichotomous coding scheme to when coding whether the script could be executed whether the results could be reproduced ('yes' or 'no'). Code often needed minor adjustments to run on the data, such as changing file locations, or loading packages in R, and coders sometimes took different approaches to how much they would adjust the code to make it run on the data. Furthermore, coders initially used different thresholds of reproducibility, based on whether every single result reported in a paper could be reproduced by the code and data, or only the main results reported in the article. After evaluating our initial coding round, we considered an article reproducible if we could get the same main results as represented in the paper with at best minor changes to make the scripts run on the data. This means we considered the analysis reproducible, even if variables could only be found with reasonable effort or by making assumptions about how variable names mapped onto variables reported in the data analysis. Furthermore, we changed folder locations when needed and loaded libraries that were required. Finally, even though sometimes Figures contained relevant information (e.g., the pattern of means) we did not require all figures to be reproducible.
